### 2019-10-09 README.Rmd
Initial commit of cleaned up regional effects repo

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# The Regional Effects of Marine Protected Areas

This repo contains code and data needed to reproduce "The Regional Effects of Marine Protected Areas" by Ovando et al. In Prep.

Users will need to install the simulation model that can be found [here](https://github.com/DanOvando/spasm), and installed using `remotes::install_github("danovando/spasm")`

>> Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000. The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030. It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders. Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies. We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection. We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA. We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region. To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.