### 2020-02-20 documents/regional-mpa-notebook.Rmd
passing to squid post empirical density ratio estimation

---
title: "regional-mpa-notebook"
author: "Dan"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)