### 2020-03-24 Untitled/Untitled.Rmd
MAJOR restructure of paper, bayesian DiD up and running, added in "ability to estiamte" test, paper 80% restructured

This PNAS journal template is provided to help you write your work in the
correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word
"Introduction" as a section heading (except for math articles in the
Physical Sciences section); this heading is implied in the first
paragraphs.

Guide to using this template {.unnumbered}
========================================

Please note that whilst this template provides a preview of the typeset
manuscript for submission, to help in this preparation, it will not
necessarily be the final publication layout. For more detailed
information please see the [PNAS Information for
Authors](http://www.pnas.org/site/authors/format.xhtml).

Author Affiliations {#author-affiliations .unnumbered}
-------------------

Include department, institution, and complete address, with the
ZIP/postal code, for each author. Use lower case letters to match
authors with institutions, as shown in the example. Authors with an
ORCID ID may supply this information at submission.

Submitting Manuscripts {#submitting-manuscripts .unnumbered}
----------------------

All authors must submit their articles at
[PNAScentral](http://www.pnascentral.org/cgi-bin/main.plex). If you are
using Overleaf to write your article, you can use the "Submit to PNAS"
option in the top bar of the editor window.

Format {#format .unnumbered}
------

Many authors find it useful to organize their manuscripts with the
following order of sections; Title, Author Affiliation, Keywords,
Abstract, Significance Statement, Results, Discussion, Materials and
methods, Acknowledgments, and References. Other orders and headings are
permitted.

Manuscript Length {#manuscript-length .unnumbered}
-----------------

PNAS generally uses a two-column format averaging 67 characters,
including spaces, per line. The maximum length of a Direct Submission
research article is six pages and a PNAS PLUS research article is ten
pages including all text, spaces, and the number of characters displaced
by figures, tables, and equations. When submitting tables, figures,
and/or equations in addition to text, keep the text for your manuscript
under 39,000 characters (including spaces) for Direct Submissions and
72,000 characters (including spaces) for

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak