### 2020-03-24 documents/Untitled.Rmd

---
title: The Regional Effects of Marine Protected Areas
author:
  - name: Daniel Ovando
    affiliation: a,1
    email: danovan@uw.edu
  - name: Jennifer E. Caselle
    affiliation: b
  - name: Christopher Costello
    affiliation: b   
  - name: Olivier Deschenes
    affiliation: b  
  - name: Steven D. Gaines
    affiliation: b
  - name: Ray Hilborn
    affiliation: a
  - name: Owen Liu
    affiliation: b        
address:
  - code: a
    address: University of Washington, School of Aquatic and Fishery Sciences
  - code: b
    address: University of California, Santa Barbara
corresponding_author:
  - code: 1
    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"
lead_author_surname: Ovando
author_contributions: |
  D.O., S.G., and R.H. developed structure of simulation model. D.O., O.D., C.C., and J.C. developed estimation strategy. J.C. provided support on collecting and interpreting data. All simulations, statistics, and sensitivity analyses performed by D.O. and O.L. . All authors contributed to the manuscript. 
conflict_of_interest: |
  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies. All of these can be interpreted as a conflict of interest when evaluating fisheries policy. C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security. Funding for D.O. was partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship.
abstract: |
  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000. The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030. It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders. The conservation effects of MPAs are most commonly evaluated using response ratios---measures of biomass densities inside their borders, relative to biomass densities in reference sites outside their borders. Studies of this nature have provided broad evidence that MPAs produce conservation benefits within their borders. While these gains can be vitally important, marine populations are rarely contained entirely within the borders of MPAs; therefore a critical question is not only do MPAs produce conservation gains inside their borders, but how do they affect the broader region in which they are located?  The Channel Islands National Marine Sanctuary provides a clear example of this challenge. A network of MPAs covering roughly 20% of the Islands’ waters was put in place in 2003, with a goal of providing regional conservation and fishery benefits. Response ratios from the region indicate that the Channel Island MPAs have increased biomass densities inside the MPAs. However, we are unable to find a clear effect of these same MPAs at the regional scale across multiple species.Building off of existing theory, we use a bio-economic simulation model to explain this discrepancy, and demonstrate under what conditions we likely can and cannot clearly estimate the regional conservation effects of MPAs. We show that MPA networks covering 25% or less of a region are likely to produce regional increases in fish biomass on the order of 10%-20%, a meaningful effect but also one that can easily be overwhelmed by environmental shocks. Our results provide a novel assessment of the regional effects of a large and iconic Marine Protected Area network, and provide guidance for managers charged with monitoring and adapting MPAs. 
significance: |
  Healthy marine ecosystems are critical to the well-being of the planet. Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems. Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes. Using simulations, we show that the regional effects of marine protected areas can be highly variable and dependent on human drivers. We find no statistically clear effect of a network of MPAs after 13 years of protection, which our simulation analysis suggests is actually to be expected. Our results lay a foundation for future research on the design and monitoring of marine protected areas. 
acknowledgements: |
  Funding for the simulation model was provided by the NMFS-Sea Grant Population and Ecosystem Dynamics Fellowship. This study would not be possible without the work provided by PISCO divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support. 
keywords:
  - Marine Protected Areas
  - Conservation
  - Bio-economic modeling
  - Program Evaluation
## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle
bibliography: My Library.bib
csl: pnas.csl
lineno: true
header-includes:
   - \usepackage{eso-pic,graphicx,transparent,xcolor}
output:
  rticles::pnas_article: default
  bookdown::word_document2: default
  bookdown::pdf_document2: default
  bookdown::html_document2: default
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
linkcolor: blue
---

<!-- \AddToShipoutPictureFG{ -->
<!--   \AtPageCenter{% or \AtTextCenter -->
<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
<!--     }} -->
<!--   } -->
<!-- } -->