### 2020-03-24 documents/Untitled.tex

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{eso-pic,graphicx,transparent,xcolor}

\templatetype{pnasresearcharticle}  % Choose template

\title{The Regional Effects of Marine Protected Areas}

\author[a,1]{Daniel Ovando}
\author[b]{Jennifer E. Caselle}
\author[b]{Christopher Costello}
\author[b]{Olivier Deschenes}
\author[b]{Steven D. Gaines}
\author[a]{Ray Hilborn}
\author[b]{Owen Liu}

  \affil[a]{University of Washington, School of Aquatic and Fishery Sciences}
  \affil[b]{University of California, Santa Barbara}

% Please give the surname of the lead author for the running footer
\leadauthor{Ovando}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Healthy marine ecosystems are critical to the well-being of the planet.
Marine protected areas, parts of the oceans protected from human
activities such as fishing, are increasingly being used in an effort to
conserve and manage these ecosystems. Our study pairs theory with
empirical methods to examine the regional effects of MPAs on
conservation outcomes. Using simulations, we show that the regional
effects of marine protected areas can be highly variable and dependent
on human drivers. We find no statistically clear effect of a network of
MPAs after 13 years of protection, which our simulation analysis
suggests is actually to be expected. Our results lay a foundation for
future research on the design and monitoring of marine protected areas.}

\authorcontributions{D.O., S.G., and R.H. developed structure of simulation model. D.O.,
O.D., C.C., and J.C. developed estimation strategy. J.C. provided
support on collecting and interpreting data. All simulations,
statistics, and sensitivity analyses performed by D.O. and O.L. . All
authors contributed to the manuscript.}

\authordeclaration{R.H's research program receives funding from environmental NGOs,
foundations, fishing industry, governments and international agencies.
All of these can be interpreted as a conflict of interest when
evaluating fisheries policy. C.C., S.G., and O.D's research program
includes funding from environmental NGOs and foundations with an
interest in ocean conservation and food security. Funding for D.O. was
partly provided by the NMFS-Sea Grant Population and Ecosystem Dynamics
Fellowship.}

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  Marine Protected Areas |  Conservation |  Bio-economic modeling |  Program Evaluation  }

\begin{abstract}
Marine Protected Areas (MPAs) cover between 3-7\% of the world's oceans,
up from less than 1\% in the year 2000. The Convention on Biological
Diversity calls for 10\% of coastal waters to be protected inside MPAs
by 2020, while the International Union for Conservation of Nature calls
for 30\% protection by 2030. It is often clear that MPAs produce
conservation benefits inside their borders, but many MPAs are also
justified on the grounds that they also benefit the broader region
outside their borders. The conservation effects of MPAs are most
commonly evaluated using response ratios---measures of biomass densities
inside their borders, relative to biomass densities in reference sites
outside their borders. Studies of this nature have provided broad
evidence that MPAs produce conservation benefits within their borders.
While these gains can be vitally important, marine populations are
rarely contained entirely within the borders of MPAs; therefore a
critical question is not only do MPAs produce conservation gains inside
their borders, but how do they affect the broader region in which they
are located? The Channel Islands National Marine Sanctuary provides a
clear example of this challenge. A network of MPAs covering roughly 20\%
of the Islands' waters was put in place in 2003, with a goal of
providing regional conservation and fishery benefits. Response ratios
from the region indicate that the Channel Island MPAs have increased
biomass densities inside the MPAs. However, we are unable to find a
clear effect of these same MPAs at the regional scale across multiple
species.Building off of existing theory, we use a bio-economic
simulation model to explain this discrepancy, and demonstrate under what
conditions we likely can and cannot clearly estimate the regional
conservation effects of MPAs. We show that MPA networks covering 25\% or
less of a region are likely to produce regional increases in fish
biomass on the order of 10\%-20\%, a meaningful effect but also one that
can easily be overwhelmed by environmental shocks. Our results provide a
novel assessment of the regional effects of a large and iconic Marine
Protected Area network, and provide guidance for managers charged with
monitoring and adapting MPAs.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{Funding for the simulation model was provided by the NMFS-Sea Grant
Population and Ecosystem Dynamics Fellowship. This study would not be
possible without the work provided by PISCO divers over the years. DO
thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments
and technical support.}

% Bibliography
% \bibliography{pnas-sample}

\end{document}