### 2020-03-27 documents/Untitled.Rmd
version sent to coauthors

---
title: The Regional Effects of Marine Protected Areas
author: Daniel Ovando, Jennifer E. Caselle, Christopher Costello, Olivier Deschenes, Steven D. Gaines, Ray Hilborn, Owen Liu
bibliography: My Library.bib
csl: pnas.csl
test: |
  thing
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2: default
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
linkcolor: blue
toc: true
---

# hello

$$d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B_cX_i} + B_s\mathbf{S_i}},shape, scale)$$