### 2020-04-26 README.Rmd
resubmissted version to PNAS (#1)
* version re-submited to PNAS

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# The Regional Effects of Marine Protected Areas

This repo contains code and data needed to reproduce "The Regional Effects of Marine Protected Areas" by Ovando et al. In Prep.

Users will need to install the simulation model that can be found [here](https://github.com/DanOvando/spasm), and installed using `remotes::install_github("danovando/spasm")`

>>  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international commitments call for 10%, 30%, and even 50% coverage. The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations. While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders. This paper examines the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and shows that even in cases when the conservation benefits are large, they are inherently difficult to detect empirically. A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits. Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level. MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present), and the size of many MPA networks suggests that effect sizes may often be smaller than 30%. Our results provide a novel assessment of the population-wide effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.