### 2020-07-18 documents/ovando-et-al-pop-effects-of-mpas.Rmd
version submitted to conservation biology

---
title: Assessing the Population-level Conservation Effects of Marine Protected Areas
abstract: |
   Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international commitments call for 10%, 30%, and even 50% coverage. The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations. While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders. We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically. A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits. Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level. We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); the size of many MPA networks suggests that effect sizes may often be smaller than 30%. Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
bibliography: references.bib
csl: conservation-biology.csl
output:
  bookdown::pdf_document2:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::html_document2: default
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
linkcolor: blue
toc: false
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
always_allow_html: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300,
                      cache = FALSE, fig.width = 3.42, fig.asp = 1.5, dev = "pdf")

```

```{r, include=FALSE}
library(sf)
library(ggmap)
library(viridis)
library(hrbrthemes)
library(scales)
library(patchwork)
library(rpart)
library(extrafont)
library(caret)
library(ggsci)
library(rEDM)
library(tidyverse)
library(ggtext)
library(here)
extrafont::loadfonts()
functions <- list.files(here::here("functions"))

walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
save_plots <- TRUE

# sim_years <- 50x
# 
# burn_years <- 20
# 
# num_patches <- 25

# run_name <- "v5.0"

width_1col <- 3.42 # inches

asp_1col <- 1.25

width_2col <- 7 # inches

asp_2col <-  .75

dpi <- 900

theme_1col <- hrbrthemes::theme_ipsum(base_size = 11,
                                      axis_title_size = 12)

theme_2col <- hrbrthemes::theme_ipsum(base_size = 12,
                                        axis_title_size = 14)

run_name <- params$run_name

sim_years <- params$sim_years

burn_years <- params$burn_years

num_patches <- params$num_patches

run_dir <- here::here("results", run_name)

experiment_dir <- here::here("results",run_name, "experiments")

load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))

load(file.path(run_dir, 'abundance_data.Rdata'))

model_runs <- read_rds( file.path(run_dir,"did_fits.rds"))

# load(file = here::here("results",run_name, "model_runs.Rdata"))

processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))

load(file = file.path(run_dir ,"sim_grid.Rdata"))

load(file = file.path(run_dir,"plots.RData"))

valgrid <- read_rds(file.path(run_dir, "valgrid.rds"))

cdfw_catches <-
  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
  group_by(sci_name, year) %>%
  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
  left_join(
    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
    by = c("sci_name" = "taxa")
  ) %>%
  filter(!is.na(classcode))

# cdfw_catches %>%
#   group_by(commonname, classcode, sci_name) %>%
#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
#   View()

density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))

outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))

outcomes <- outcomes %>% 
  filter(!is.na(mpa_effect))

density_ratios <- density_ratios %>% 
  filter(!is.na(biased_density_ratio))

   eqo <- outcomes %>%
     filter(year == max(year))

facet_labels <- c(
  mpa_size = "Range in MPA",
  depletion = "Depletion"
)

testplot <- ggplot()

panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")

panel_width = unit(1,"npc") - sum(ggplotGrob(testplot)[["widths"]][-3]) - unit(4,"line")

gc <- guide_colorbar(frame.colour = "black",
                     ticks.colour = "white",
                     barheight = panel_height)

  hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 10)

    wide_hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 25)

theme_set(theme_1col)

# models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)

# model_runs <- model_runs %>%
#   filter(models_worked) %>%
#   mutate(tmb_fit = map(tmb_fit,"result")) %>%
#   mutate(processed_fits = map(tmb_fit, process_fits)) %>%
#   mutate(did_plot = map(processed_fits, "did_plot"))

base_run <- model_runs %>%
filter(var_names == "pisco_a", data_to_use == "all", center_scale == TRUE)

mpa_run <- model_runs %>%
  filter(data_source == "pisco",
         var_names == "pisco_a",
         data_to_use == "mpa_only")

fishsed_run <- model_runs %>%
  filter(data_source == "pisco",
         var_names == "pisco_a",
         data_to_use == "fished_only")

kfm_run <- model_runs %>%
  filter(data_source == "kfm")

used_data <- abundance_data$data[abundance_data$data_source == "pisco"][[1]]

tmdata <- used_data %>% 
  filter(targeted == 1) %>% 
  select(tm, everything())

median_tm = round(median(tmdata$tm))

site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
  magrittr::set_colnames(., tolower(colnames(.))) %>%
  select(
    site,
    side,
    mpagroup,
    mpa_status,
    reserve,
    region,
    year_mpa,
    mpaareanm2,
    lat_wgs84,
    lon_wgs84
  ) %>%
  rename(lat_wgs84 = lon_wgs84,
         lon_wgs84 = lat_wgs84) %>%
  unique() %>%
  mutate(eventual_mpa = (year_mpa > 0))

top_species <- used_data$classcode %>% unique()

life_summaries <- life_history_data %>% 
  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
  unique() %>% 
  filter(classcode %in% top_species) %>% 
  select(-classcode) %>% 
  gather(variable,value) %>% 
  group_by(variable) %>% 
  summarise(min_val = min(value, na.rm = TRUE),
         max_val = max(value, na.rm = TRUE))

 sim_life_history <- sim_grid %>%
    mutate(
      age_mature = map_dbl(fish, "age_mature"),
      vbk = map_dbl(fish, "vbk"),
      linf = map_dbl(fish, "linf")
    ) %>%
    select(scientific_name, age_mature, vbk, linf) %>%
    unique() %>% 
   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
          linf > life_summaries$min_val[life_summaries$variable == "linf"],
          linf < life_summaries$max_val[life_summaries$variable == "linf"])

 year_bins <-
  seq(2003, max(pisco_data$year) + 1, by = 3)

 sim_mpa_effects <- outcomes %>% 
  filter(set == "ci",
         final_depletion > 0) %>% 
  mutate(year = years_protected + 2002) %>% 
  filter(year <= max(pisco_data$year)) %>% 
  mutate(binned_year = cut(year, year_bins, include.lowest = FALSE))  %>% 
  filter(year >= 2003,
         !is.na(binned_year)) %>% 
  mutate(mpa_effect = pmin(mpa_effect,3))

 did_fit <- model_runs$did_fit[model_runs$data_source == "pisco" &
                                 model_runs$var_names == "pisco_a" &
                                 model_runs$data_to_use == "all"][[1]]$did_reg

 if (!dir.exists(here("documents", "figs"))){

   dir.create(here("documents", "figs"))

 }
```

<!-- \AddToShipoutPictureFG{ -->
<!--   \AtPageCenter{% or \AtTextCenter -->
<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
<!--     }} -->
<!--   } -->
<!-- } -->

# Introduction {#intro .unnumbered}

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources. Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978]. Modern MPAs were first established as marine analogs to the terrestrial protection of iconic landscapes [@iucn1976]. Over time our goals and expectations for MPAs have evolved and become more ambitious. Recent international agreements to expand MPAs are based on the assumption that well-designed MPAs will achieve benefits both within, and outside, their borders[@gaines2010].

The empirical MPA literature has focused on assessing the ability of MPAs to provide conservation gains within their borders [@lester2009; @edgar2014; @gerber2005]. As conservation benefits accrue inside MPAs, MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort. These dynamics result in MPAs affecting not only the organisms within their borders, but the broader connected population of affected species. Theory states that numerous factors influence how MPAs might affect fish populations. These include the scale of adult and larval dispersal relative to the size of the MPAs [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008;@gerber2005], the strength, timing, and location of density dependence [@burgess2014], the degree of enforcement [@edgar2014], how overfished the population would be without the MPA, the time span under evaluation, and how fishing and management responds to the implementation of the MPAs [@hastings2003;@gerber2003; @hilborn2004a; @gaines2010; @botsford2003;@smith2003a;@smith2006a;@walters2000a;@white2011,;@white2010c;@ovando2016;@moffitt2013]. The location and spacing of the MPAs within a network can also influence population-level impacts through network effects [@gaines2010; @costello2010a].

What empirical evidence is there for these population-level effects of MPAs? Conservation outcomes of MPAs are generally evaluated by "response ratios", commonly measured as the ratio of biomass densities of species inside relative to outside of MPAs [@lester2009; @halpern2003a;@edgar2014;@caselle2015]. These studies found clear evidence that well enforced and sufficiently sized MPAs are associated with high response ratios. Several studies have also documented empirical evidence for the existence of adult or larval fish spillover affecting fish abundance [@halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009]. MPAs have also caused alterations in fishing effort [@costa2013;@mason2012;@murawski2005;@mcdermott2019]. Why does this large body of empirical work not serve as sufficient evidence for the population-level effects of MPAs?

Spillover-based studies provide evidence that MPAs may have effects outside their borders, but they are not sufficient evidence for the population effect of MPAs. Response ratios have shortcomings as measures of population-level effects as well. Control sites used in calculating response ratios are often selected based on habitat characteristics [@ferraro2018]. However, export of adults or larvae from the MPA to these "control" site affects their status as controls, as does displacement of fishing effort from MPAs to control sites. In theory, control sites far enough away to negate both biological spillover and concentration by the fishing fleet could be selected, but finding suitably far sites that are also appropriate proxies for the ecological and economic context of the MPAs is challenging. Response ratios can be a highly imprecise and biased measure of the population-level conservation effect of an MPA network [@ferraro2018;@moffitt2013]. As response ratios are our primary source of evidence for the conservation effects of MPAs, this means that our collective understanding of the population-level impacts of MPAs is surprisingly limited.

<!-- Given the potential unreliability of response ratios for the task, how can we empirically measure the population-level effects of MPAs? The perfect experiment would involve two parallel worlds that were identical, except for the implementation of an MPA.  In world “A”, no MPA would be implemented, and in the facsimile world “B”, the MPA would be implemented. Both worlds would be tracked before and after MPA implementation, and the biomass densities would be compared after treatment. Instead of two parallel worlds, a similar experimental design would involve random placement of MPAs.  Unfortunately, neither of these experimental designs has been implemented, because parallel worlds do not exist and MPAs are to our knowledge never placed at random. Careful causal inference requires some other means of controlling for biases introduced by factors such as the unobserved environmental shocks, the MPA siting process, biological spillover, and concentration of fishing effort  [@ferraro2018;@larsen2019]. -->

<!-- As stakeholders around the world seek to use MPAs in marine resource management portfolios, it is critical that we develop a better understanding of the magnitude and drivers of population-level MPA effects. This is particularly true in the context of a changing climate, in which "set it and forget it" management strategies are increasingly untenable; we need to be able to monitor the performance of MPAs in order to adapt them as dictated by a shifting environment. From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA? From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area? Overall, what are the population-level effects of MPAs?  -->

The MPAs within the Channel Islands National Marine Sanctuary, California, USA (hereafter the Channel Islands) provide an opportunity to improve our understanding of population-level MPA effects. A network of protected areas covering approximately 20% of the Islands' waters was put in place in 2003 [@kirlin2013]. The network has been used as a model case study in protected area design around the world (@botsford2014 and references therein). We use data from the first 14 years of protection to provide what is to our knowledge the first empirical assessment of the population-level effect of a large MPA network on a wide array of fin-fish species. In contrast to clear differences in biomass densities observed inside and outside of well-protected MPAs both globally [@lester2009] and in the Channel Islands [@caselle2015] we are unable to detect a clear population effect from the Channel Islands MPAs. We build off of existing MPA theory to consider why this might be, and provide guidance for scientists and managers as to when and how we might expect to estimate the population-level conservation effects of MPAs.

# Methods {#methods .unnumbered}

Our primary methods consist of a spatially-explicit bio-economic simulation model and a Bayesian difference-in-difference regression. We use the simulation model to provide provide theoretical context for our empirical results. The regression provides our estimate of the population-level conservation effect of the Channel Islands MPAs. All analysis were conducted in R [@rcoreteam2019]. Our difference-in-difference model was fit using Stan [@carpenter2017] using the `rstanarm` package [@goodrich2020]. All data and code needed to fully reproduce this manuscript are publicly available at [github.com/DanOvando/population-effects-of-mpas](https://github.com/DanOvando/population-effects-of-mpas). A detailed description of the simulation model structure, as well as sensitivity analyses of our estimation model, are available in the online supporting information (SI).

## Simulation Model {#simmodel .unnumbered}

Our bio-economic model simulates the effect of MPAs on a spatially explicit age-structured representation of a fish population. Readers can explore the functionality of the model using an online tool available at [danovando.shinyapps.io/simmpa/](https://danovando.shinyapps.io/simmpa/).

The simulation model consists of 50 patches with wrapped edges. For any one simulation we randomly pull a species and its associated life history traits from the `FishLife` [@thorson2017c] package. We pair these data with randomly selected values governing the characteristics of the simulation (Table.\@ref(tab:sim-table)). The population begins at unfished equilibrium. After fishing effort is applied, each simulation is run to equilibrium with and without the selected MPA design (holding all else constant). In each time step, we measure the difference in biomass densities between the scenario with and without the MPAs to calculate the population effect of the MPAs over time. Simulation results provide a library of plausible MPA effects for a range of biological and economic assumptions.One set of simulations is specifically designed to reflect the dynamics of the Channel Islands. We only include species of the same genus as those targeted by fishing in the Channel Islands, restrict fishing pressure to be mostly moderate to low, and cap the MPA size at 20% of the population's range (a likely upper end to the effective MPA size in question, @rassweiler2012).

## Difference in Difference Regression {#did .unnumbered}

The difference-in-difference model used empirical kelp forest survey data from the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) monitoring in the Northern Channel Islands. PISCO conducts visual SCUBA surveys at a large number of rocky reef and kelp forest sites inside and outside of MPAs throughout the Channel Islands, producing estimates of densities of fishes that are both targeted and non-targeted by fishing (Fig.\ref{fig:ci-map}). The details of the monitoring program are described in @caselle2015. We define the population-level conservation effects of MPAs as the change in mean total biomass densities of targeted fin-fish both inside and outside of MPAs, relative to the mean total biomass densities of targeted fin-fish inside and outside of MPAs that would have occurred without the MPAs.

Building off of the concepts explored in @caselle2015, we used an identification strategy utilizing biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 0])` species that are not directly targeted by fishing as our control group (non-targeted), and biomass densities of `r n_distinct(used_data$classcode[used_data$targeted == 1])` species targeted by fishing as our treatment group. Targeted species in the Channel Islands include fin-fish such as California sheephead (*Semicossyphus pulcher*), and copper (*Sebastes caurinus*) and blue (*Sebastes mystinus*) rockfish. Non-targeted species include garibaldi (*Hypsypops rubicundus*), halfmoons (*Medialuna californiensis*), and blacksmith (*Chromis punctipinnis*). We used a Bayesian difference-in-difference regression to estimate any difference in mean total biomass densities of fin-fish species targeted by fishing effort (i.e., those potentially affected by an MPA) and those species not targeted by fishing before and after MPA implementation. This identification strategy controls for unobserved environmental shocks to the system that are independent of the MPAs. Conditional on the assumptions of the model, this regression produces an estimate of the effect of the MPAs on the mean total biomass densities of targeted species throughout the Channel Islands. For example, consider an evenly distributed population that has 50% of its range protected by an MPA. If the MPA increases biomass densities inside its boundaries by 20%, and by 0% outside the reserve, the population effect of the MPA would be 10%.

 <!-- Our analysis omits species from important invertebrate fisheries including red urchin (*Mesocentrotus franciscanus*) and spiny lobster (*Panulirus interruptus*), although our methods could similarly be applied to these species given data.  -->

<!-- The regression amounts to estimating the pre-post MPA difference in the mean total biomass density of targeted species minus the same difference for non-targeted species in the Channel Islands. -->

The simplified form of this model is

\begin{equation}
  d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape, scale)
\label{eq:did}
\end{equation}

where $d_i$ is the biomass density at observation *i*, *T* indicates whether the observation *i* is for a targeted ($T = 1$) or non-targeted ($T = 0$) species, and *MPA* marks whether observation *i* is in a pre MPA ($MPA = 0$) or post MPA ($MPA = 1$) state. **B^c^** is a vector of coefficients for additional control variables in matrix *X* such as water visibility and observer experience. **B^s^** is a vector of hierarchical coefficients for each sampling location **S** , clustered by island.  Under the assumptions of this model, $\beta_3$ is the causal effect of the treatment (*MPA*) on the treated (targeted species). $\beta_0$ is the mean total biomass of non-targeted species pre-MPA, and $\beta_1T$ is the mean total biomass of targeted species pre-MPA. $\beta_2MPA$ is the effect of the post-MPA period on the non-targeted species. The shape and scale parameters of the Gamma distribution are estimated as well (Table.\@ref(tab:coef-table)).

<!-- r percent(mean(used_data$eventual_mpa))` of the samples in the raw data come from within MPAs, while the MPAs themselves cover 20% of the surface waters of the Channel Islands. As such, we weight each observation such that within MPA data are assigned a total weight of 0.2 and outside-MPA data a total weight of 0.8. The intercept prior was was determined using the `rstanarm` `autoscale` function. For the non-intercept terms, we manually set a prior as a normal distribution with a mean of 0 and a standard deviation of 2. This implies that coefficients such as the MPA effect have a prior that provides support for an effect size centered on zero with a range of roughly -4 to 4. Since the covariates are centered and scaled, and the dependent variable is on the log-scale, this is an extremely diffuse prior (e.g. a one standard deviation change in an independent variable changes the dependent variable by 4 on the log scale) -->

# Results {#results .unnumbered}

Our Channel Islands-specific simulation provides a set of outcomes that can feasibly bound our expectations for the empirical effects of the MPA network. The targeted species in our database span a range of life histories, but are largely made up of fishes in the perch and rockfish complexes, with a mean Von Bertalanffy growth coefficient of `r round(mean(used_data$k[used_data$targeted == 1]),2)`, and a median age at maturity of `r round(median(used_data$tm[used_data$targeted == 1]),0)` years, and a mix of exploitation histories, ranging from low to high (but mostly relatively low) [@alonzo2004;@dick2017]. While a wide range of MPA effects are plausible, from 0% to upwards of 150%, the median simulated effect size for this stylized version of the Channel Islands after over ten years of protection was `r percent(median(sim_mpa_effects$mpa_effect[sim_mpa_effects$years_protected > 10]))` (Fig.\@ref(fig:mpa-effect)).

We compared this range of plausible outcomes to the empirical data. @caselle2015 found a statistically significant increase in the response ratios of targeted species over time, and evidence that this increase is smaller for non-targeted species. Updating the results of @caselle2015 through 2017 shows a continuation of the increasing trend in the response ratios of targeted species (Fig.\@ref(fig:rr-plot)). This provides evidence that the Channel Islands MPAs are large enough and sufficiently well-enforced as to provide meaningful protection within their borders.

These response ratios cannot, however, be used as a definitive indicator of population effects of the MPAs. In the case of the Channel Islands MPAs, control sites are often located within a few kilometers of an MPA, suggesting that they are susceptible to both local biological spillover and by concentration of fishing effort excluded from the MPAs. To illustrate this problem, we paired our simulated Channel Islands response ratios to the estimated posterior probability distributions of the empirical Channel Island response ratios by matching the patterns in the empirical response ratios. The response ratio trends we observe in the data could plausibly be produced by a wide range of population-level MPA effects (Fig.\@ref(fig:rr-plot)). Response ratios well over one were associated with "true" population-level MPA effects generally less than 25%, and, importantly, many simulations produced large response ratios but population-level MPA effects close to 0%. This can occur if for example fishing pressure is only moderate, adult movement is low, larval dispersal is high, and displaced fishing effort concentrates around the border of the MPAs.

<!-- For example, if the mean estimated response ratio in the year 2006 (year three after implementation) is one, we found all simulations that had response ratios near one after three years of protection, and then pulled out the "true" simulated population-level MPA effect from each of those simulations. This provides us with a distribution of simulated population-level MPA effects that could plausibly generate the types of response ratios actually observed in the Channel Islands. -->

```{r}
did_model <- model_runs %>%
  filter(data_source == "pisco",
         var_names == "pisco_a",
         data_to_use == "all") %>% {
           .$did_fit[[1]]$did_reg
         }
did_coefs <- did_model %>% broom::tidy(intervals = TRUE,conf.level = 0.9)

```

Our empirical difference-in-difference regression provides a different perspective on population-wide MPA effects. Over the first three years of implementation (2003-2006), the effects of the MPAs are unclear, with support for a small negative effect to a substantially positive effect, with much higher probability of a small positive effect (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`,
90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2003,2006]"]) - 1, accuracy = 1)`). Over the next six years the model estimates greater probabilities of an increasingly positive MPA effect, peaking in 2009-2011 with a median estimate of MPA effect of a `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` increase in mean total biomass density of targeted species (90% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2009,2012]"]) - 1, accuracy = 1)` (Fig.\@ref(fig:mpa-effect)). These empirical estimates are in line with the outcomes that our simulation model suggests are plausible. However, in the subsequent years the trend reverses itself, and in 2015-2017 we once again see no clear effect of the MPAs (median estimated effect `r percent(exp(did_coefs$estimate[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`,
90\% credible interval `r percent(exp(did_coefs$lower[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`
- `r percent(exp(did_coefs$upper[did_coefs$term == "targeted:year_bins(2015,2018]"]) - 1, accuracy = 1)`.) Fig.\@ref(fig:mpa-effect).

# Discussion {#discussion .unnumbered}

Containing a carefully designed, well-enforced, and well-studied MPA network, the Channel Islands seems to be an ideal location to study the population-level effects of protected areas. The persistently high response ratios suggest that despite overall decreases in targeted biomass densities inside and outside MPAs relative to pre-MPA measurements, the MPAs may still be providing protection within their borders. But, these response ratios are not necessarily an indicator of broader population-level effects. The difference-in-difference strategy presents an alternative identification strategy. It is not without its own strict caveats, but provides some potential improvements over response-ratios as a means of estimating population-level effects. While we estimate an uncertain but overall positive effect of the MPA network in its first few years, we are unable to detect a robust positive signal from 2012-2017. After 14 years of MPA protection we are left without a clear picture of the effect of the Channel Island MPA network on biomass densities of targeted fin-fish species from either response ratios or the difference-in-difference model.

How can we explain this result? Much of the theoretical literature on MPAs assumes that larger reserves produce larger conservation gains [@white2011 and references therein]. However, these models generally simulate fleet dynamics through fishing mortality rates (e.g. concentration of fishing mortality, @halpern2004). The implication of this is that total catches will scale with the size of the available population.

```{r}
positive_outcomes <- outcomes %>% 
  filter(years_protected == 15) %>% 
  group_by(fleet_model) %>% 
  summarise(pp = mean(mpa_effect > 0))

```

Alternatively, fishers can pursue a  “constant-catch” strategy, where fishers have a catch objective and exert as much (or little) effort as needed to achieve that objective. Subsistence fisheries may use a constant-catch style policy over the short-term, as they seek to ensure that their food needs are met. Constant-catch dynamics might also occur in fisheries with constraining quotas that are not updated after the implementation of MPAs. Fishers pursuing constant catch in areas outside an MPA have to fish harder to achieve the same catch from a smaller part of the population, causing a population conservation loss under `r percent(1 - positive_outcomes$pp[positive_outcomes$fleet_model == "constant-catch"], accuracy = 1)` of our constant-catch simulations. This is an important and often overlooked possibility, especially as MPAs are increasingly implemented in quota-managed fisheries [@liu2018].

While we do not have access to fine scale fishing data from the Channel Islands alone, reported catches for the species of interest in the Santa Barbara region exhibit a mix of stable, downward, and upward trajectories (SI Fig.S1) which indicates that a negative MPA effect caused by a constant-catch fishing strategy is unlikely. However, some of the most common targeted species in the Channel Islands, such as blue rockfish and copper rockfish, have seen dramatic increases in reported catches in the Santa Barbara region since the year 2010 (although other such as California sheephead have seen declines relative to pre-MPA catch level). While we include catch histories in our regression, to the extent that these increases in catches are due to external market forces, the disappearance of clear positive population-level MPA effects estimated by our model could be due to exogenous increases in fishing pressure overwhelming any population-level increases caused by the MPA network.

Another possibility for the estimated decline in Channel Islands MPA effects is environmental disturbance. The Channel Islands region experienced a dramatic ‘marine heatwave’ beginning in 2014 and persisting through 2016, resulting in part in extremely elevated water temperatures throughout the region [@gentemann2017]. Many of the non-targeted species in the Channel Islands have warm thermal affinities and have increased in numbers since the heatwave [@freedman2019]. The targeted group is made up mostly of fishes with cold-water affinities. We hypothesize that the recent evidence for a decline in densities of targeted species is primarily due to environmental conditions that disproportionately affected the targeted group. The practical result of this is that in the presence of this marine heatwave the non-targeted species may no longer serve as an effective control for the evolution of biomass densities of targeted fin-fish in the absence of the MPAs, given the magnitude of the environmental shock relative to the size of the population-level MPA effect.

<!-- This hypothesis is supported by the fact that we estimate a similar decline in the estimated MPA effect both inside and outside of the MPAs (SI Appendix Fig.S18) (as well as observe the decline in the raw data, Fig.\@ref(fig:raw-density)). If the cause of recent declines was due to increases in fishing pressure we would expect to see substantial declines only in the fished areas. While we see evidence for the validity of the parallel trends assumption in the early years of our data (Fig.\@ref(fig:mpa-effect), SI Appendix Fig.S26), it is likely that the non-targeted species reacted in a different way to the marine heatwave than the targeted species did, causing a violation of the parallel trends assumption in the post 2014 data. -->

<!-- Simulation analysis suggests that we should not be surprised by this result. The Channel Islands MPAs cover 20% of the surface waters in the Channel Islands, and while formal stock assessments are not available for many of the targeted species in our analysis, what evidence we have suggests that, as a group, these fish are on average not heavily overfished. Some species, such as California sheephead and blue rockfish were likely below target levels during the period [@alonzo2004;@dick2017], but projections based upon the overall average response across all species will likely suggest modest benefits even if a subset of species could experience much larger population gains. Under such circumstances, the average percentage difference in densities of targeted species with and without MPAs is likely to be modest (Fig.\ref{fig:mpa-effect}). Effects of this size will often be challenging to detect empirically given the large natural variation of marine ecosystems (especially temperate reefs) and the observation error inherent in infrequent (e.g. annual) common in marine systems.  -->

<!-- Our simulations and difference-in-difference model examine percentage changes in biomass densities (as opposed to changes in total population size). An increase in biomass densities from 0.02kg/m^2^ to 0.04kg/m^2^ would be measured as a 100% increase. While a large percentage effect, this is a small change in biomass densities relative to the variance in the observation process itself. While an MPA providing 25% protection might be expected to produce a 30% increase in biomass densities, expressed as a percentage of unfished biomass, this gain is on the order of 15% (SI Appendix Fig.S45). In addition, the median estimated age at sexual maturity for the targeted species included in this study is `r median_tm` years, meaning that the span of this analysis represents less than three generations of MPA protection for half of the measured species. Ongoing monitoring may yet reveal clearer effects, especially if the time series extends beyond the periods of environmental shocks. Analysis of more rapidly growing and maturing species, e.g. spiny lobster, may also reveal clearer signals. -->

Trophic cascades are another possible explanation for our equivocal results. All of the species in this empirical analysis may affect each other through mechanisms such as predation, competition, and habitat modification. We used convergent cross mapping (CCM), in the manner of @clark2015, to test for significant dynamic interactions between species and therefore the possibility of the trophic cascades biasing our results. We found no significant cross-mappings between targeted and non-targeted species, indicating that while clearly there are interactions between these groups on some level, the effects within the timespan of the data are not pronounced enough to be of concern to our results (SI Fig.S42). However, the longer MPAs are in place, the greater the possibility that substantial species interactions that can affect use of non-targeted species as a control may arise.

Our finding of small MPA effects in recent years is therefore likely a result of both changing fishing dynamics and environmental disturbance. Given the natural variability of marine ecosystems, how large of an effect would an MPA network have to have in order to allow a difference-in-difference strategy such as this to be a reliable measure of MPA effects? Our library of simulation results allows us to explore how MPA effects, and our ability to measure them, might vary with observation error and natural fish recruitment variation (including the possibility of recruitment regimes). Using a Bayesian difference-in-difference regression on these simulated data, we estimated the percent error between the posterior probability distribution of the estimated MPA effect from the regression and the true simulated MPA effect.

The model struggled severely when MPA effect sizes were less than 25% and the model was faced with both observation and process errors (Fig.\@ref(fig:val-plot)). Even models fit to data generated from large effect sizes commonly mis-estimated the true MPA effect by 50% or more.  Obtaining a mean absolute percent error (MAPE) of 25% or less across our simulated datasets required a population-level MPA effect of at least 30%. In the context of the Channel Islands, this finding suggests that we should not be surprised at our difficulty in precisely estimating the population-level effect of the MPAs.  
<!-- This is merely an illustrative exercise, omitting critical factors such as detection probability and sampling strategy. However, since nearly any omission which one can think of would make an MPA effect harder to detect, not easier, these results serve as a useful floor for the likely difficulty in estimating MPA effects. -->

Two of the most critical drivers of MPA effect size are the size of the MPA network and the degree of fishing pressure. Looking across these two variables, based on our simulations the MPA network must be large (25% or more of a species range) and the target species overfished (pre-MPA depletion greater than about 60%) to achieve an effect size with a likely MAPE of 25% or less. Although recently some extremely large MPAs have been enacted that may indeed reach into the higher levels of MPA coverage, most MPA networks for near-shore commercial fin-fish are likely to cover areas more in line with the Channel Islands (20%) or smaller. As such, many MPA networks are expected to have population-level effect sizes that are difficult to detect unless target species were experiencing extreme overfishing prior to MPA establishment (Fig.\@ref(fig:sim-effects)-A).

<!-- The MPA literature highlights a large number of variables beyond simply size and fishing pressure that can affect performance. We examined the variability in expected MPA effects across each of the two major axes (pre-MPA depletion and MPA size). As both MPA size and pre-MPA depletion increase, the potential for a large MPA effect increases. However, for both variables even at extremely large or extremely small values a wide range of MPA effects were possible (though as we might expect the effect of pre-MPA depletion, in other words fishing pressure, was much clearer than MPA size alone) (Fig.\@ref(fig:sim-effects)-B).  -->

<!-- MPAs are an important part of the marine resource management toolbox. Under ideal circumstances they can protect individual species and ecosystem linkages, while supporting local economies through tourism and fishing opportunities. One rationale for the expansion of MPAs is that they will deliver net conservation benefits both inside *and outside* their borders. To date, this assumption is insufficiently tested, and this is the focus of our paper. Our results show that population-level conservation benefits of MPAs are highly context dependent and in many circumstances, are likely to be small enough that they are nearly impossible to detect empirically. This is exactly what we found in our empirical case study from the relatively large and well-studied network of MPAs in the Channel Islands, California, USA.  -->

<!-- A large theoretical literature has pointed out the importance of a wide range of economic and environmental factors in determining the effectiveness of MPAs [e.g. @hilborn2004; @smith2006a; @white2010c]. Our simulation model allows users to place this collective theory into context, clarifying under what states of the world we might expect to see particular effect sizes, and demonstrating what effect sizes we might reasonably expect to be detectable in the real world. In the case of the Channel Islands, simulation modeling suggests for the network in question, an effect size on the order of 30% or smaller is likely, which our analysis in turn demonstrates may be challenging to detect even with a comprehensive monitoring program. -->

<!-- Once simulation models have been used to help design an MPA, how can users evaluate whether it is achieving their objectives?  -->
@lester2009 found massively higher densities and biomass inside MPAs than outside [median biomass increases of  over 400%]. But, when MPAs affect nearby control sites used in response ratios through biological spillover or concentration of fishing effort, it is entirely possible to for MPA to produce massive response ratios while simultaneously having minimal effects on the entire population partially protected by the MPAs (and *vice versa*).

As @larsen2019 suggests, there are many potential alternatives for estimating the population effects of MPAs that better account for the challenges of causal inference (though that may be more data-intensive). We applied one such approach here, and yet were still unable to reach robust conclusions as to the effect of MPAs on the total biomass density of targeted fin-fish in the Channel Islands, due to the likely small size of the true effect relative to the influence of environmental variability.

While this does not mean that all MPAs will face similar challenges in estimating their effects, our results in the relatively large, well-enforced, and rigorously studied (though also relatively lightly fished) Channel Islands Marine Protected Area network make clear that in many instances empirically detecting a clear effect of MPAs on total biomass densities of targeted fin-fish at the population level may not be possible. In the absence of clear empirical findings, simulation modeling can help inform the range of effect sizes that may be expected, and monitoring programs can be tuned to focus on the species groups that have the highest chance of a detectable effect size over the early years of the reserve. Expanding data collection to include robust monitoring of spatio-temporal fleet dynamics may help assess the validity of control sites used in response ratios, support the direct inclusion of these fleet dynamics into statistical models, and allow managers to take into account potential negative interactions between MPAs and fleet dynamics such as those that may occur under constant-catch dynamics. Whenever possible monitoring programs should be implemented prior to MPA implementation to provide a pre-treatment benchmark.

Stock assessment models that account for the population within the MPAs [@field2006a] may be able to answer the management relevant question of whether fishing mortality rates and biomass levels from systems containing MPAs are in-line with management objectives [e.g. @nickols2019]. However, such an approach does not necessarily shed light on whether the MPAs themselves caused the estimated state of the population, and of course are highly data intensive, potentially restricting our ability to provide stock-assessment based inference of MPA outcomes for broad arrays of targeted species.

Non-equilibrium analyses also help set expectations for effect sizes over time (@nickols2019;@kaplan2019;@ovando2016;Fig.\ref{fig:mpa-effect}). Educating communities about the challenges of estimating the effects of MPAs can help ensure that a lack of a clear effect is not necessarily viewed as a failure of the program, or large positive result based solely on response ratios as a clear sign of success and subsequent relaxation of other fishery management strategies. Rather, results and subsequent management actions must be considered in the context of reasonable expectations given the size, age, and degree of enforcement of the MPAs in question, together with the ecological and economic dynamics of a given system. While this paper has focused on the conservation outcomes of MPAs, future work must also address the challenge of predicting and estimating the fishery impacts of protected areas, an important question for which extremely little empirical evidence is available [@smith2006a].

MPAs are an important part of the marine resource management toolbox. As the number and size of global MPA networks increase, we must set appropriate expectations for their outcomes and design effective monitoring programs. While the history of MPA science has made important strides in helping us understand the dynamics of protected areas, the future of MPA science must directly tackle the challenge of evaluating the performance of these MPAs at the regional scale. This is particularly true if communities are depending on MPAs as their primary marine resource management tool. Commonly employed metrics such as response ratios may be applicable in some circumstances, but are vulnerable to inaccuracy or misuse as metrics of population-level effects. Bio-economic modeling can help frame community expectations, reducing the potential for a reduction in support if unrealistic conservation or fishery expectations are not realized. Statistical approaches like difference-in-difference regression that explicitly address complications such as the spatial spillover effects of MPAs may give users an improved understanding of the performance of their MPAs, but even they may struggle when expected effect sizes are small. Clearly communicating what we should expect, and what we can detect, from MPAs is critical to ensuring that MPAs play effective roles in fisheries management and marine conservation.

<!-- Dependence on unreliable estimators of MPA effects may lead to stakeholders incorrectly attributing negative environmental shocks as MPA failures, or interpreting data arising from scorched-earth fishing outside MPAs as a conservation success. Both of these scenarios would hinder the ability of MPAs to serve as effective marine resource management tools at scale. -->

# Literature Cited {#lit .unnumbered}

<div id="refs"></div>

# Tables {#tables .unnumbered}

```{r sim-table, results = "asis", echo=FALSE}
variable <- c(
"Scientific Name",
"steepness",
"Adult movement ($\\sigma_{s=a}$)",
"Larval movement ($\\sigma_{s=l}$)",
"Recruitment variation ($\\sigma_{r}$)",
"Recruitment autocorrelation ($ac_{r}$)",
"DD adult movement",
"Density-dependence timing ",
"% Patches in MPA",
"Initial fishing relative to natural mortality",
"Selectivity as a multiple of maturity length",
"Fleet model",
"Spatial effort model",
"Years into simulation to start MPA",
"MPA is Larval Source?",
"Randomly place MPA?",
"Fleet reaction to MPA",
"Patchiness",
"MPA habitat factor")

distribution <- 
c(
"Drawn from all possible species in `FishLife` (@thorson2017c)",
"uniform(0.6,0.95)",
"uniform(0,0.25 * P)",
"uniform(0,0.25 * P)",
 "$\\in{0,0.05,.1,.2}$",
 "$\\in{0,0.05,.1,.2}$",
 "$\\in{0.25,1}$",
 "$\\in{Local, Global,PostDispersal}$",
 "uniform(0.01,1)",
 "uniform(0.01,4)",
 "uniform(0.1,1.25)",
"$\\in{OpenAccess, ConstantEffort,ConstantCatch}$",
 "$\\in{Uniform, Biomass,Profits}$",
 "uniform(5,0.66)",
"$\\in{TRUE,FALSE}$",
"$\\in{TRUE,FALSE}$",
"$\\in{Concentrate,Leave}$",
"uniform(0.01,0.75)",
"$\\in{1, 4}$"
)

equations <- tibble(Variable = variable, Distribution = distribution )

knitr::kable(equations,
             escape = FALSE,
             caption = "Variables for MPA simulations",
             format = "pipe") %>% 
  kableExtra::kable_styling()

```

```{r coef-table, results = "asis", echo=FALSE}

coefs <- did_fit %>%
  broom::tidy(intervals = TRUE, prob = 0.89) %>%
  select(term, estimate, lower, upper) %>%
  rename(
    Mean = estimate,
    `Lower 89th Percentile` = lower,
    `Upper 89th Percentile` = upper
  )

  knitr::kable(
    coefs,
    digits = 2,
    caption = glue::glue("Posterior means and credible interval for key model coefficients. Conditional on the assumptions of the model, the targeted:year_bins coefficients reflect the effect of the MPAs on mean total biomass of targeted finfish in the Channel Islands. targeted indicates whether the species in question is targeted, tex = technican experience, surge is mean reported surge, kelp is mean reported kelp cover, lag catch is the total reported catch for targeted species in the previous year, temp is the mean temperature at a site, temp_dev is the deviation of a temperature observation from the mean temperatuer for that island in question. Raw data contained {nrow(used_data)} individual observations, total number of targeted and non-targeted mean mean total biomass densities by site and year used to fit model was {nrow(did_fit$data)}"),
    format = "pipe")

```

# Figures {#figures .unnumbered}

```{r ci-map, include = TRUE,fig.width=width_1col, fig.asp=1.5, fig.cap="Map of study region; the Northern Channel Islands, California, USA. Colors show binned number of PISCO sampling events across the time period of our study.", include=TRUE}

temp_plot <- sample_location_plot + 
  theme(axis.text = element_blank())

temp_plot
ggsave(
  temp_plot,
  filename = here::here("documents","figs","sample_location_plot.pdf"),
  width = width_1col,
  height = width_1col * asp_1col,
  device = cairo_pdf
)

```

```{r raw-density, fig.width = width_1col, fig.asp=asp_1col,fig.cap = "Centered and scaled trends in biomass densities of targeted and non-targeted fin-fish included in our study Top panel shows trends across all sites, with smaller background lines showing trends for each individual species. Bottom two panels show aggregate biomass density trends outside and inside MPAs"}

temp_plot <-
  (total_trend_plot / (mpa_trend_plot)) & theme(legend.position = "top", plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "lines"),
                                                panel.grid = element_blank()
  )

temp_plot

ggsave(
  temp_plot,
  filename = here::here("documents","figs","raw_biomass_density_plot.pdf"),
  width = width_1col,
  height = width_1col * asp_1col,
  device = cairo_pdf
)

```

```{r rr-plot, include = TRUE, fig.width=width_1col, fig.asp=asp_1col,,fig.cap="90% Posterior probability distributions of response ratios for targeted species (x-axis) over time (y-axis) in dark grey. Simulated population-level (pop.) effect on biomass densities matched to empirical response ratios in light grey. For response ratios, a value of zero indicates that biomass densities of targeted species are identical inside and outside MPAs, a value of one that biomass densities of targeted species are 100% greater inside MPAs relative to outside. For MPA population effect, a value of zero indicates that biomass densities are identical in the with- and without- MPA scenarios. A value of 1 indicates that biomass densities are 100% greater in the scenario with MPAs than the scenario without MPAs."}

temp_plot <- response_ratio_plot + 
  theme_1col + 
  theme(axis.title.x = element_textbox(size = 10)) + 
  geom_vline(aes(xintercept = 0), color = "black", linetype = 2)

ggsave(
  temp_plot,
  filename = here::here("documents","figs","response_ratio_plot.pdf"),
  device = cairo_pdf,
  width = width_1col,
  height = width_1col * asp_1col,
  units = "in",
)

temp_plot
# pdf(file =  "figs/density-ratio.png",width = 8, height = 5, useDingbats = TRUE)
# print(dr_plot)
# dev.off()

```

```{r mpa-effect, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Results of difference-in-difference regression estimating the population effect of the Channel Island MPAs on mean total biomass densities of targeted species (difference in mean total biomass density of targeted species over time relative to expected levels using non-targeted species as a control). Grey distributions show posterior probability distribution of estimated MPA effect; red point is median estimated effect, thicker red section 50% credible interval, thinner red line 90% credible interval). Blue distributions in background show range of MPA effects produced by simulation model tuned to reflect the dynamics of the Channel Island MPAs (black dashed line is median simulated value). Results are estimated in blocks of three years, including years greater than or equal to left-hand value and less than right-hand value."}

mpa_effect_plot

ggsave(
  mpa_effect_plot,
  filename = here::here("documents","figs","mpa_effect_plot.pdf"),
  width = width_2col,
  height = width_2col * asp_2col,
  device = cairo_pdf
)

```

```{r val-plot, include = TRUE,fig.width = width_1col, fig.asp=asp_1col, fig.cap = "Distribution of percent error in posterior estimates of population-level MPA effect (y-axis) plotted against true simulated MPA effect (x-axis). Shading shows concentration of simulations. Black line shows mean absolute percent error (MAPE) as a function of true simulated MPA effect."}

temp_plot <- validation_plot + 
  theme_1col + 
  theme(legend.position = "top") #+ 
  # scale_fill_viridis(option = "C",
  #                    name = "Sims(#)",
  #                    guide = hgc <- guide_colorbar(frame.colour = "black",
  #                                                  ticks.colour = "black",
  #                                                  barwidth = ggplot2::unit(5,"lines")))

temp_plot
ggsave(
  temp_plot,
  filename = here::here("documents","figs","validation_plot.pdf"),
  width = width_1col,
  height = width_1col * asp_1col,
  device = cairo_pdf
)

```

```{r sim-effects, include=TRUE, fig.width=width_2col, fig.asp=asp_2col, fig.cap = "Simulated population-level (pop.) MPA effect sizes as a function of percent of species' range inside MPA (x-axis), and pre-MPA depletion (y-axis). Pre-MPA depletion is a measure of fishing pressure, where 0 means that the population is unfished, and 1 means that the population is extinct in the time period immediately prior to MPA implementation. Panel A) shows median MPA effects across range in MPA and pre-MPA depletion Panel B) shows distribution of simulations across range of MPA size and pre-MPA depletion separately."}

expected_mpa_effect_plot
ggsave(
  expected_mpa_effect_plot,
  filename = here::here("documents","figs","expected_mpa_effect.pdf"),
  width = width_2col,
  height = width_2col * asp_2col,
  device = cairo_pdf

)

```