### 2020-12-02 documents/test-tables.Rmd
fixed that damn conservation-letters.csl error

---
title: Assessing the Population-Level Conservation Effects of Marine Protected Areas
abstract: |
  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, up from less than 1% in the year 2000, and international organizations call for 30% by 2030. The premise underlying MPA expansion is that they conserve biodiversity, habitats, and fished populations. While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders. We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically. A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits. Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach. We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are more common) are even harder to detect. Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area network, and provide guidance for communities charged with monitoring and adapting MPAs.
bibliography: references.yaml
csl: conservation-biology.csl
output:
  bookdown::html_document2: default
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: xelatex
params:
  run_name: ["v6.0"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
linkcolor: blue
toc: false
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{draftwatermark}
  - \SetWatermarkText{DRAFT}
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(flextable)
```

Hmmm [@hastings2003;@gerber2003;@botsford2003;@smith2003a;@hilborn2004a;@gaines2010;@white2011;@moffitt2013;@ovando2016;@walters2000a]