### 2020-12-07 README.Rmd
updating README with renv instructions

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Assessing the Population-level Conservation Effects of Marine Protected Areas

This repo contains code and data needed to reproduce "Assessing the Population-level Conservation Effects of Marine Protected Areas" by Ovando et al. In Prep.

The script `make-pop-effects-of-mpas.R` will reproduce all results and manuscripts for this project.

This project is set up with [`renv`](https://rstudio.github.io/renv/articles/renv.html) to manage package dependencies. Inside R (and with your working directory set correctly) run `renv::restore()`. Follow all prompts. This will install the correct versions of all the packages needed to replicate our results. Packages are installed in a stand-alone project library for this paper, and will not affect your installed R packages anywhere else.

>>  Marine Protected Areas (MPAs) cover 3-7% of the world’s ocean, and international organizations call for 30% by 2030. While numerous studies show that MPAs produce conservation benefits *inside* their borders, many MPAs are also justified on the grounds that they confer conservation benefits to the broader population *beyond* their borders. We examine the conditions under which MPAs can provide population-level conservation benefits inside and outside their borders, and show that even in cases where the population benefits are large, they are inherently difficult to detect empirically. A network of MPAs was put in place in The Channel Islands National Marine Sanctuary in 2003, with a goal of providing regional conservation and fishery benefits. Evidence indicates that the Channel Island MPAs have increased biomass densities inside the MPAs, but we are unable to find a clear effect of these same MPAs at the population level using a Bayesian difference-in-difference approach. We show that MPA effect sizes less than 30% are likely to be difficult to detect (even when they are present); smaller effect sizes (which we find are common) are even harder to detect. Our results provide a novel assessment of the population-level effects of a large and iconic Marine Protected Area