### 2021-05-02 documents/pop-effects-of-mpas-front-matter.Rmd
pre-print cleanup

---
output:
  bookdown::word_document2:
    reference_docx: template.docx
  pdf_document: default
  word_document: default
---

## Title  {.unnumbered}
Assessing the Population-level Conservation Effects of Marine Protected Areas

## Article Impact Statement {.unnumbered}

The population-level conservation effects of Marine Protected Areas are likely to be extremely challenging to measure and detect

## Running Head {.unnumbered}

The Population-Level Effects of MPAS

## Keywords: {.unnumbered}

Marine Protected Area Networks

Marine Conservation

Program Evaluation

Causal Inference

Bio-economic Modeling

Channel Islands National Marine Sanctuary

## Word Count {.unnumbered}

5994

## Author Details {.unnumbered}

Daniel Ovando*
School of Aquatic and Fishery Sciences
University of Washington, Seattle, WA 98195-5020 USA
danovan@uw.edu

Jennifer E. Caselle
Marine Science Institute, University of California Santa Barbara, Santa Barbara, CA, USA

Christopher Costello
Environmental Market Solutions Lab, University of California, Santa Barbara, CA, USA

Olivier Deschenes
Environmental Market Solutions Lab, University of California, Santa Barbara, CA, USA 
Department of Economics, University of California, Santa Barbara, CA, USA

Steven D. Gaines
Bren School of Environmental Science and Management, University of California, Santa Barbara, CA 93117

Ray Hilborn
School of Aquatic and Fishery Sciences
University of Washington, Seattle, WA 98195-5020 USA

Owen Liu
Bren School of Environmental Science and Management, University of California, Santa Barbara, CA 93117

## Corresponding Author  {.unnumbered}

Daniel Ovando
Fishery Science Building
University of Washington
1122 NE Boat Street, Seattle, WA 98195-5020 USA
danovan@uw.edu

## Acknowledgements  {.unnumbered}

Funding for the simulation model was provided by the National Marine Fisheries Service-SeaGrant Population and Ecosystem Dynamics Fellowship. Empirical data collection was funded primarily by the David and Lucille Packard Foundation in support of the Partnership for Interdisciplinary Studies of Coastal Oceans (PISCO) with additional funding from the California Ocean Protection Council and California SeaGrant. This study would not have been possible without the work provided by PISCO divers over the years. We especially thank K. Davis Koehn, and A. Parsons-Field. We also thank C. Szuwalski, J. Lawson, A. Punt, C. Brown, and 2 anonymous reviewers for helpful comments and technical support.